A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-06 latex/main.tex
Updates from Overleaf

% We used more seeds, but in practise it doesn't matter and it's just
% complicating the story.
For Delta
(first detected in December, 2020 in India \cite{ECDCVOCUpdates}),
% More samples were unconditionally added,
% but the earliest are the seeds.
we used two seeds:
one sample (ERR5461562) assigned to B.1.617.1 (Kappa) and
the other (ERR5676810) assigned to B.1.617.2 (Delta).
Kappa is closely related to Delta, and
therefore seeding it helps to reconstruct the origin of Delta.
% See https://github.com/jeromekelleher/sc2ts-paper/blob/main/inference/viridian_config.toml
Both of these seeds were sampled in the UK,
the Kappa one in February, 2021 and the Delta one in March, 2021.
% https://github.com/jeromekelleher/sc2ts-paper/issues/267
% The reality is a bit more complicated,
% but this is easy to explain.
These are the earliest Kappa and Delta samples
in the Viridian v04 dataset
that were also used for Pango designation.
% FIXME more Delta details and some references to back up our assertions
% There's some specific info in the Viridian paper on this, I'm sure.
% We should mention long branch splitting here too.]
The resulting subgraph around Delta in the ARG
can be seen in Figure~\ref{fig:subgraph_delta};
for further explanation, see Document~\ref{sec:dd_delta}.

@@

[
FIXME: Discuss  the large effect recombinants inferred by both:
sc2ts identifies two recombination events with \textgreater 10k descendants (specifically, 166,979 and 166,285 descendant samples), both of which we identify as spurious (see \ref{sec:spurious-ancestrally-important-recombinants}). RIPPLES at $p=4$ also identifies two recombination events with \textgreater 10k descendants (531,224 and 339,523 descendant samples); these do not correspond to the same events as in sc2ts. At $p=3$, RIPPLES identifies six recombination events with \textgreater 10k descendant samples, four of which have \textgreater 100k descendant samples.
]

>>>>

% We used more seeds, but in practise it doesn't matter and it's just
% complicating the story.
For Delta
(first detected in December, 2020 in India \cite{ECDCVOCUpdates}),
% https://github.com/jeromekelleher/sc2ts-paper/issues/267
% More samples were unconditionally added,
% but the earliest are the seeds.
we used two seeds:
% https://github.com/jeromekelleher/sc2ts-paper/blob/main/inference/viridian_config.toml
ERR5461562 assigned to B.1.617.1/Kappa
(dated February 22, 2021 from the UK) and
ERR5676810 assigned to B.1.617.2/Delta
(dated March 23, 2021 from the UK).
Kappa is closely related to Delta \cite{Stern2021}, and
therefore seeding it helps to reconstruct the origin of Delta.
% What we actually did is a bit more complicated,
% as it was manual, but this is easier to explain.
These are the earliest Kappa and Delta samples
in the Viridian v04 dataset
% https://github.com/jeromekelleher/sc2ts-paper/issues/267
% SHZ: I'd like to recheck this.
that were used for the designation of these Pango lineages.
% FIXME more Delta details and some references to back up our assertions
% There's some specific info in the Viridian paper on this, I'm sure.
% We should mention long branch splitting here too.]
The resulting subgraph around Delta in the ARG
can be seen in Figure~\ref{fig:subgraph_delta};
for further explanation, see Document~\ref{sec:dd_delta}.

@@

[
FIXME: Discuss the large effect recombinants inferred by both in more detail. E.g.:
sc2ts identifies two recombination events with \textgreater 10k descendants (specifically, 166,979 and 166,285 descendant samples),
both of which we identify as spurious (see \ref{sec:spurious-ancestrally-important-recombinants}).
RIPPLES at $p=4$ also identifies two recombination events with \textgreater 10k descendants (531,224 and 339,523 descendant samples);
these do not correspond to the same two spurious large-effect events in sc2ts. At $p=3$, RIPPLES identifies six recombination events with \textgreater 10k descendant samples, four of which have \textgreater 100k descendant samples.
]
